# Results

## Exisitence of Gender Inequalities in Education 
```{r,echo=FALSE}
data_new<- read.csv(file = 'data_new.csv')
library(tidyr)
library(ggplot2)
library(ggridges)
library(tidyverse)
```

```{r}
myData <- as.data.frame(data_new)
lr_a = subset(myData,lr_a!= "NA",lr_y!="NA")
noNA_1 <- pivot_longer(lr_a, cols=lr_a:lr_y, names_to = "lr", values_to = "lr_rate")
ggplot(noNA_1, aes(x = lr, y = lr_rate, fill = Sex))+
  geom_boxplot(color="#88ada6",
                 alpha=.75,
                 bins = 15)+
  labs(title="Literacy rate and gender",
       x = 'Literacy rate in differnet period',
       y = 'Literacy rate')+
  theme(plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        axis.text = element_text(size=9),
        axis.title = element_text(size=14, face="bold"))
```
    
From this boxplot, we can know the variability or dispersion of the data. the literacy rate in youth is similar in both genders. The 50% of lr_y(the literacy rate in youth) in both genders are between 90% and 100%, and the median in males and females is around 100%. To be more specific, the literacy in females are slightly higher than in male, respectively 99% and 98%. There are outliers, and the minimum rate in females is 29% and 49% for males. For adults, half of lr_a(the literacy rate in adults) in different sex have different dispersion. The rates in females have larger distributions and outliers. In females, most of the rates are between 71% and 96%, while most of the data in the male are higher than in females, and 50% of male rates are between 82% and 98%. The median in males(95%) is slightly higher than 2% in females(93%). 

In short, the literacy rates in youth have higher literacy rates than the literacy rate in adults in both genders. It illustrates that education has been improved a lot in both genders. In the past, literacy rates in males are higher than in females, and now literacy rates in females have better performance than in males. However, there are many countries in outliers that have a low literacy rate. We still need to improve education in those countries. We also will figure out if this situation is related to the developing regions.



```{r}
cr = subset(myData,cr_p!= "NA",cr_ls!="NA")
noNA_2 <- pivot_longer(cr, cols=cr_p:cr_us, names_to = "cr_type", values_to = "c_rate")
ggplot(noNA_2, aes(x = cr_type, y = c_rate, fill = Sex))+
  geom_boxplot(color="#88ada6",
                 alpha=.75,
                 bins = 15)+
  scale_x_discrete(limits = c("cr_us","cr_ls","cr_p"))+
  labs(title="compelete school rate and gender",
       x = 'compelete school rate in differnet period',
       y = 'compelete school rate')+
  theme(plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        axis.text = element_text(size=9),
        axis.title = element_text(size=14, face="bold"))+
  coord_flip()
```  

The second box plot is related to the complete school rate in primary school, lower secondary, and upper secondary. In general, the complete school rate in primary school is higher than in lower secondary and then upper secondary. The maximum rates of females in each period are higher than males, but the minimum rates are lower than males. In primary school, the distribution of rates in each sex is quite similar in the interquartile range, and the median in females is higher than in males. In lower secondary, the complete school rates in females become dispersive and interquartile ranges are larger than in males. There is the same trend in the upper secondary, and the median female complete school rate is lower than males. 

In short, some countries achieve higher rates, around 100%, but the trend of complete school rate from young ages to teenage ages is decreasing. The complete school rate in females have more dispersive.

    
    
```{r}
os = subset(myData,os_p!= "NA",os_us!="NA")
noNA_3 <- pivot_longer(os, cols=os_p:os_us, names_to = "os_type", values_to = "os_rate")
ggplot(noNA_3, aes(x = os_rate, y = os_type, fill = Sex))+
  geom_density_ridges(alpha = 0.6) +
  theme_ridges() +
  scale_y_discrete(limits = c("os_us","os_ls","os_p"))+
  labs(title="out of school rate and gender",
       x = 'out of school rate',
       y = 'out of school rate in differnet period')+
  theme(plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        axis.text = element_text(size=9),
        axis.title = element_text(size=14, face="bold"))
```      
We use the ridgeline to show the out-of-school rate in different genders. The ridgeline shows the distribution of out-of-school values for 2 groups. All graphs in primary school, lower secondary, and upper secondary show right skew, which is a good sign, and the distribution of out-of-school rates focuses on the small rates. In primary school and lower secondary, they have a parallel distribution in both genders. The out-of-school rate in lower secondary slightly move to the right. The female rates in lower secondary are higher than male rates in the peak, and then the male rates are higher than females. For upper secondary, it is clear to observe that the distribution moves to the right a lot, and the rates at peak become larger around 20%.

## Possibility of Gender Inequality for Certain Countries in Terms of educational underachievement 


```{r,fig.width=8,fig.height=12,echo=FALSE}
ggplot(data_new%>%filter(!is.na(comp_higher_2yrs_2529_m)), aes(comp_higher_2yrs_2529_m,reorder(country,comp_higher_2yrs_2529_m))) +
        geom_line(aes(group = country)) +
        geom_point(aes(color = Sex))+facet_grid(Development_Regions~ .,scales="free_y") +ggtitle("Tertiary completion rate vs Sex")+theme(axis.text=element_text(size=10))
```

```{r,fig.width=8,fig.height=12,echo=FALSE}
ggplot(data_new%>%filter(!is.na(attend_higher_1822_m)), aes(attend_higher_1822_m,reorder(country,attend_higher_1822_m))) +
        geom_line(aes(group = country)) +
        geom_point(aes(color = Sex))+ggtitle("Higher education attendance rate vs Sex")+theme(axis.text=element_text(size=10))
```

```{r,fig.width=8,fig.height=12,echo=FALSE}
ggplot(data_new%>%filter(!is.na(eduyears_2024_m)), aes(reorder(country,eduyears_2024_m),eduyears_2024_m,fill=Sex)) +
  geom_bar(position = "dodge", stat = "identity",width=0.7)+
  theme(axis.text=element_text(size=10))+coord_flip()
```

## Relationship among Development Level, Education Underachievement and Gender Inequality




